Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ \smaller \textbf{{Berlin version \YellowPaperVersionNumber}}

### 2021-11-26 Paper.tex
Replace SHA3 opcode with KECCAK256 (#825)

\nopagebreak
\begin{tabu}{l r l}
\toprule
Name & Value & Description \\
\midrule
$G_{\mathrm{zero}}$ & 0 & Nothing paid for operations of the set {\small $W_{\mathrm{zero}}$}. \\
$G_{\mathrm{jumpdest}}$ & 1 & Amount of gas to pay for a {\small JUMPDEST} operation. \\
$G_{\mathrm{base}}$ & 2 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{base}}$}. \\
$G_{\mathrm{verylow}}$ & 3 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{verylow}}$}. \\
$G_{\mathrm{low}}$ & 5 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{low}}$}. \\
$G_{\mathrm{mid}}$ & 8 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{mid}}$}. \\
$G_{\mathrm{high}}$ & 10 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{high}}$}. \\
$G_{\mathrm{warmaccess}}$ & 100 & Cost of a warm account or storage access. \\
$G_{\mathrm{coldaccountaccess}}$ & 2600 & Cost of a cold account access. \\
$G_{\mathrm{coldsload}}$ & 2100 & Cost of a cold storage access. \\
$G_{\mathrm{sset}}$ & 20000 & Paid for an {\small SSTORE} operation when the storage value is set to non-zero from zero. \\
$G_{\mathrm{sreset}}$ & 2900 & Paid for an {\small SSTORE} operation when the storage value's zeroness remains unchanged or\\
&&is set to zero. \\
$R_{\mathrm{sclear}}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from\\
&&non-zero.\\
\linkdest{R__selfdestruct}{}$R_{\mathrm{selfdestruct}}$ & 24000 & Refund given (added into refund counter) for self-destructing an account. \\
\linkdest{G__selfdestruct}{}$G_{\mathrm{selfdestruct}}$ & 5000 & Amount of gas to pay for a {\small SELFDESTRUCT} operation. \\
$G_{\mathrm{create}}$ & 32000 & Paid for a {\small CREATE} operation. \\
\linkdest{G__codedeposit}{}$G_{\mathrm{codedeposit}}$ & 200 & Paid per byte for a {\small CREATE} operation to succeed in placing code into state. \\
$G_{\mathrm{callvalue}}$ & 9000 & Paid for a non-zero value transfer as part of the {\small CALL} operation. \\
\linkdest{G__callstipend}{}$G_{\mathrm{callstipend}}$ & 2300 & A stipend for the called contract subtracted from $G_{\mathrm{callvalue}}$ for a non-zero value transfer. \\
\linkdest{G__newaccount}{}$G_{\mathrm{newaccount}}$ & 25000 & Paid for a {\small CALL} or {\small SELFDESTRUCT} operation which creates an account. \\
$G_{\mathrm{exp}}$ & 10 & Partial payment for an {\small EXP} operation. \\
$G_{\mathrm{expbyte}}$ & 50 & Partial payment when multiplied by the number of bytes in the exponent for the {\small EXP} operation. \\
$G_{\mathrm{memory}}$ & 3 & Paid for every additional word when expanding memory. \\
\linkdest{G__txcreate}{}$G_\text{txcreate}$ & 32000 & Paid by all contract-creating transactions after the {\textit{Homestead} transition}.\\
\linkdest{G__txdatazero}{}$G_{\mathrm{txdatazero}}$ & 4 & Paid for every zero byte of data or code for a transaction. \\
\linkdest{G__txdatanonzero}{}$G_{\mathrm{txdatanonzero}}$ & 16 & Paid for every non-zero byte of data or code for a transaction. \\
\linkdest{G__transaction}{}$G_{\mathrm{transaction}}$ & 21000 & Paid for every transaction. \\
$G_{\mathrm{log}}$ & 375 & Partial payment for a {\small LOG} operation. \\
$G_{\mathrm{logdata}}$ & 8 & Paid for each byte in a {\small LOG} operation's data. \\
$G_{\mathrm{logtopic}}$ & 375 & Paid for each topic of a {\small LOG} operation. \\
$G_{\mathrm{sha3}}$ & 30 & Paid for each {\small SHA3} operation. \\
$G_{\mathrm{sha3word}}$ & 6 & Paid for each word (rounded up) for input data to a {\small SHA3} operation. \\
$G_{\mathrm{copy}}$ & 3 & Partial payment for {\small *COPY} operations, multiplied by words copied, rounded up. \\
$G_{\mathrm{blockhash}}$ & 20 & Payment for {\small BLOCKHASH} operation. \\

@@

C_{\mathrm{aaccess}}(\boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}, A) + G_{\mathrm{copy}}\times\lceil\boldsymbol{\mu}_{\mathbf{s}}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
C_{\mathrm{aaccess}}(\boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}, A) & \text{if} \quad w \in W_{\mathrm{extaccount}}\\
G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1] & \text{if} \quad w = \text{\small LOG0} \\
G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG1} \\
G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+2G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG2} \\
G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+3G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG3} \\
G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+4G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG4} \\
C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}, A) & \text{if} \quad w \in W_{\mathrm{call}} \\
C_\text{\tiny SELFDESTRUCT}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SELFDESTRUCT} \\
G_{\mathrm{create}} & \text{if} \quad w = \text{\small CREATE}\\
G_{\mathrm{create}}+G_{\mathrm{sha3word}} \times \lceil \boldsymbol{\mu}_{\mathbf{s}}[2] \div 32 \rceil & \text{if} \quad w = \text{\small \hyperlink{create2}{CREATE2}}\\
G_{\mathrm{sha3}}+G_{\mathrm{sha3word}} \times \lceil \boldsymbol{\mu}_{\mathbf{s}}[1] \div 32 \rceil & \text{if} \quad w = \text{\small SHA3}\\
G_{\mathrm{jumpdest}} & \text{if} \quad w = \text{\small JUMPDEST}\\
C_\text{\tiny SLOAD}(\boldsymbol{\mu}, A, I) & \text{if} \quad w = \text{\small SLOAD}\\
G_{\mathrm{zero}} & \text{if} \quad w \in W_{\mathrm{zero}}\\
G_{\mathrm{base}} & \text{if} \quad w \in W_{\mathrm{base}}\\
G_{\mathrm{verylow}} & \text{if} \quad w \in W_{\mathrm{verylow}}\\
G_{\mathrm{low}} & \text{if} \quad w \in W_{\mathrm{low}}\\
G_{\mathrm{mid}} & \text{if} \quad w \in W_{\mathrm{mid}}\\
G_{\mathrm{high}} & \text{if} \quad w \in W_{\mathrm{high}}\\
G_{\mathrm{blockhash}} & \text{if} \quad w = \text{\small \hyperlink{blockhash}{BLOCKHASH}}\\
\end{cases}
\end{equation}
\begin{equation}
w \equiv \begin{cases} I_{\mathbf{b}}[\boldsymbol{\mu}_{\mathrm{pc}}] & \text{if} \quad \boldsymbol{\mu}_{\mathrm{pc}} < \lVert I_{\mathbf{b}} \rVert\\
\text{\small STOP} & \text{otherwise}
\end{cases}
\end{equation}

@@

\begin{tabu}{\usetabu{opcodes}}
\toprule
\multicolumn{5}{c}{\textbf{20s: SHA3}} \vspace{5pt} \\
\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
0x20 & {\small SHA3} & 2 & 1 & Compute Keccak-256 hash. \\
&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \mathtt{KEC}(\boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] \dots (\boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1] - 1) ])$ \\
&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[1])$ \\
\bottomrule
\end{tabu}

>>>>

\nopagebreak
\begin{tabu}{l r l}
\toprule
Name & Value & Description \\
\midrule
$G_{\mathrm{zero}}$ & 0 & Nothing paid for operations of the set {\small $W_{\mathrm{zero}}$}. \\
$G_{\mathrm{jumpdest}}$ & 1 & Amount of gas to pay for a {\small JUMPDEST} operation. \\
$G_{\mathrm{base}}$ & 2 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{base}}$}. \\
$G_{\mathrm{verylow}}$ & 3 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{verylow}}$}. \\
$G_{\mathrm{low}}$ & 5 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{low}}$}. \\
$G_{\mathrm{mid}}$ & 8 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{mid}}$}. \\
$G_{\mathrm{high}}$ & 10 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{high}}$}. \\
$G_{\mathrm{warmaccess}}$ & 100 & Cost of a warm account or storage access. \\
$G_{\mathrm{coldaccountaccess}}$ & 2600 & Cost of a cold account access. \\
$G_{\mathrm{coldsload}}$ & 2100 & Cost of a cold storage access. \\
$G_{\mathrm{sset}}$ & 20000 & Paid for an {\small SSTORE} operation when the storage value is set to non-zero from zero. \\
$G_{\mathrm{sreset}}$ & 2900 & Paid for an {\small SSTORE} operation when the storage value's zeroness remains unchanged or\\
&&is set to zero. \\
$R_{\mathrm{sclear}}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from\\
&&non-zero.\\
\linkdest{R__selfdestruct}{}$R_{\mathrm{selfdestruct}}$ & 24000 & Refund given (added into refund counter) for self-destructing an account. \\
\linkdest{G__selfdestruct}{}$G_{\mathrm{selfdestruct}}$ & 5000 & Amount of gas to pay for a {\small SELFDESTRUCT} operation. \\
$G_{\mathrm{create}}$ & 32000 & Paid for a {\small CREATE} operation. \\
\linkdest{G__codedeposit}{}$G_{\mathrm{codedeposit}}$ & 200 & Paid per byte for a {\small CREATE} operation to succeed in placing code into state. \\
$G_{\mathrm{callvalue}}$ & 9000 & Paid for a non-zero value transfer as part of the {\small CALL} operation. \\
\linkdest{G__callstipend}{}$G_{\mathrm{callstipend}}$ & 2300 & A stipend for the called contract subtracted from $G_{\mathrm{callvalue}}$ for a non-zero value transfer. \\
\linkdest{G__newaccount}{}$G_{\mathrm{newaccount}}$ & 25000 & Paid for a {\small CALL} or {\small SELFDESTRUCT} operation which creates an account. \\
$G_{\mathrm{exp}}$ & 10 & Partial payment for an {\small EXP} operation. \\
$G_{\mathrm{expbyte}}$ & 50 & Partial payment when multiplied by the number of bytes in the exponent for the {\small EXP} operation. \\
$G_{\mathrm{memory}}$ & 3 & Paid for every additional word when expanding memory. \\
\linkdest{G__txcreate}{}$G_\text{txcreate}$ & 32000 & Paid by all contract-creating transactions after the {\textit{Homestead} transition}.\\
\linkdest{G__txdatazero}{}$G_{\mathrm{txdatazero}}$ & 4 & Paid for every zero byte of data or code for a transaction. \\
\linkdest{G__txdatanonzero}{}$G_{\mathrm{txdatanonzero}}$ & 16 & Paid for every non-zero byte of data or code for a transaction. \\
\linkdest{G__transaction}{}$G_{\mathrm{transaction}}$ & 21000 & Paid for every transaction. \\
$G_{\mathrm{log}}$ & 375 & Partial payment for a {\small LOG} operation. \\
$G_{\mathrm{logdata}}$ & 8 & Paid for each byte in a {\small LOG} operation's data. \\
$G_{\mathrm{logtopic}}$ & 375 & Paid for each topic of a {\small LOG} operation. \\
$G_{\mathrm{keccak256}}$ & 30 & Paid for each {\small KECCAK256} operation. \\
$G_{\mathrm{keccak256word}}$ & 6 & Paid for each word (rounded up) for input data to a {\small KECCAK256} operation. \\
$G_{\mathrm{copy}}$ & 3 & Partial payment for {\small *COPY} operations, multiplied by words copied, rounded up. \\
$G_{\mathrm{blockhash}}$ & 20 & Payment for {\small BLOCKHASH} operation. \\

@@

C_{\mathrm{aaccess}}(\boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}, A) + G_{\mathrm{copy}}\times\lceil\boldsymbol{\mu}_{\mathbf{s}}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
C_{\mathrm{aaccess}}(\boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}, A) & \text{if} \quad w \in W_{\mathrm{extaccount}}\\
G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1] & \text{if} \quad w = \text{\small LOG0} \\
G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG1} \\
G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+2G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG2} \\
G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+3G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG3} \\
G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+4G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG4} \\
C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}, A) & \text{if} \quad w \in W_{\mathrm{call}} \\
C_\text{\tiny SELFDESTRUCT}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SELFDESTRUCT} \\
G_{\mathrm{create}} & \text{if} \quad w = \text{\small CREATE}\\
G_{\mathrm{create}}+G_{\mathrm{keccak256word}} \times \lceil \boldsymbol{\mu}_{\mathbf{s}}[2] \div 32 \rceil & \text{if} \quad w = \text{\small \hyperlink{create2}{CREATE2}}\\
G_{\mathrm{keccak256}}+G_{\mathrm{keccak256word}} \times \lceil \boldsymbol{\mu}_{\mathbf{s}}[1] \div 32 \rceil & \text{if} \quad w = \text{\small KECCAK256}\\
G_{\mathrm{jumpdest}} & \text{if} \quad w = \text{\small JUMPDEST}\\
C_\text{\tiny SLOAD}(\boldsymbol{\mu}, A, I) & \text{if} \quad w = \text{\small SLOAD}\\
G_{\mathrm{zero}} & \text{if} \quad w \in W_{\mathrm{zero}}\\
G_{\mathrm{base}} & \text{if} \quad w \in W_{\mathrm{base}}\\
G_{\mathrm{verylow}} & \text{if} \quad w \in W_{\mathrm{verylow}}\\
G_{\mathrm{low}} & \text{if} \quad w \in W_{\mathrm{low}}\\
G_{\mathrm{mid}} & \text{if} \quad w \in W_{\mathrm{mid}}\\
G_{\mathrm{high}} & \text{if} \quad w \in W_{\mathrm{high}}\\
G_{\mathrm{blockhash}} & \text{if} \quad w = \text{\small \hyperlink{blockhash}{BLOCKHASH}}\\
\end{cases}
\end{equation}
\begin{equation}
w \equiv \begin{cases} I_{\mathbf{b}}[\boldsymbol{\mu}_{\mathrm{pc}}] & \text{if} \quad \boldsymbol{\mu}_{\mathrm{pc}} < \lVert I_{\mathbf{b}} \rVert\\
\text{\small STOP} & \text{otherwise}
\end{cases}
\end{equation}

@@

\begin{tabu}{\usetabu{opcodes}}
\toprule
\multicolumn{5}{c}{\textbf{20s: KECCAK256}} \vspace{5pt} \\
\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
0x20 & {\small KECCAK256} & 2 & 1 & Compute Keccak-256 hash. \\
&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \mathtt{KEC}(\boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] \dots (\boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1] - 1) ])$ \\
&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[1])$ \\
\bottomrule
\end{tabu}
